---
title: 'PHP2550: Project 1'
author: "Alitzel Serrano Laguna"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
options(knitr.kable.NA = '')
```

### Overview 

The data used in this project came from merging (see pda_project.R) study data of 
both parent and child into a dataset containing parent ID as a unique identifier.
The data had 49 total parents and included 77 total variables relating to 
socioeconomic status, parent smoking status during and after pregnancy at 
different time points, parent and baby cotinine levels, 
children substance use, and externalizing and self-regulation variables. In
summary, this report is sectioned into the following  parts: reformatting and 
missing data, baseline characteristics, relevant variables and 
socioeconomic covariates, correlation and composite scores, and aims.

### Reformatting 

Reformatting included changing character type to categorical variables, and 
numeric variables to categorical variables, and correcting data entry errors. 
These variables were reformatted to allow for easier analysis (see code appendix). 
For example, the variable **mom_numcig** was classified as character and included
entries such as "2 black and miles a day" and ""44989". The 44989 cigarettes
smoked per day was identified as an outlier and potentially a data entry error and
changed to NA.

In the initial summary of missing data by column, the reported number of substance  
use for 30-days resulted in high percentages. In further observing the data, 
NA values for those columns were then coded 0, if the corresponding 
ever used a substance variable was 0=No. 


```{r}
#Load Libraries
library(naniar)
library(tidyverse)
library(gtsummary)
library(gridExtra)
```

```{r}
#Load data
setwd("/Users/Ali23/OneDrive/Desktop/PDA/")
data <- read.csv("project1.csv")  #project.csv data created from pda_project.R
```


### Missing Data Patterns

In observing missing data, it was found that 63 variables had at least 
one missing observation. The variable **mom_smoke_pp1** with 79.59% missing data 
had by far the highest percent missing. **mom_smoke_pp2** had 40.82% missing. All
other SDP and postpartum smoking variables had less than 20% missing data. ETS **smoke_exposure**
variables had similar missing percentages ranging around 20%-23%. Outcome variables such as
**childasd**, **num_alc_30**, **num_e_cigs_30**, and **bpm_int** also had over 20% missing. 
Parental monitoring question also had a high percentage missing. Income and child's sex
were also variables with more than 25% missing data. We also see the same number of 
missing values for many related variables, potentially belonging to the same participants
who had missing data across. 

Additionally it was observed that there was substantial missing data for SDP at 16, 22, and 32 weeks and for postpartum 
(PP smoking). Knowing that cigarette smoking is an additive substance, to deal with this issue, it was assumed that if a parent smoked or did not smoke at 16 and 22 weeks, 
then the same behavior continued at **32weeks**. It was also assumed that if a parent 
smoked or did not smoke at 22 and 32 weeks and at postpartum visit 2, then the 
same behavior continued at **postpartum visit 1**. No assumptions were made for
environmental tobacco smoke (ETS) exposure variables because this was more subject to
change compared to parent smoking behavior. 

In the missing data by rows, eight parents 
had greater than 65% missing variable data. Due to the small sample size N = 49, 
these participants were kept in the analysis except one. This parent with ID: 53902 was removed as
they were missing all cotinine measurements and all SDP/ postpartum/ ETS data. If a parent had at least one category of SDP, PP, and ETS as well as at least one cotinine measure, then based on this data, it would be possible to calculate a propensity score or conduct multiple imputation
based on parents with similar information. This was not done in this report, but in further analysis of this data it 
should be considered.


Table 1 reports variables who had greater than 25% missing data.
Table 2 reports greater than 50% missing data by parent.


```{r}
#Reformatting Variables and missing data

#create NA for entries "" for mom_num_cig & SDP and post-pregnancy smoking var
data[,c(21:28)] <- lapply(data[,c(21:28)], function(x) ifelse(x == "", NA, x))

#change numeric to categorical for sex, ethnic, race, edu, employ vars
data[,c(3:13)] <- data[,c(3:13)] %>%
  mutate_if(is.numeric, factor)  # for parent
data[,c(53:61)] <- data[,c(53:61)] %>%
  mutate_if(is.numeric, factor)  # for child

#for substance use, if child answered no, set corresponding numeric value to 0
data$num_cigs_30 <- ifelse(data$cig_ever == 0, 0, data$num_cigs_30)
data$num_alc_30 <- ifelse(data$alc_ever == 0, 0, data$num_alc_30)
data$num_mj_30 <- ifelse(data$mj_ever == 0, 0, data$num_mj_30)
data$num_e_cigs_30 <- ifelse(data$e_cig_ever == 0, 0,  data$num_e_cigs_30)

## changing the mom_num_cig variable ##
data$mom_numcig <- as.character(data$mom_numcig)
#set "none" equal to 0
data$mom_numcig <- ifelse(data$mom_numcig == "None",  "0", data$mom_numcig )
data$mom_numcig <- ifelse(data$mom_numcig == "2 black and miles a day",  "2", 
                          data$mom_numcig)
#average 20-25 to 22.5
data$mom_numcig <- ifelse(data$mom_numcig == "20-25","22.5",data$mom_numcig)
#change 44989 to NA
data$mom_numcig <- ifelse(data$mom_numcig == "44989",NA,data$mom_numcig)
data$mom_numcig <- as.numeric(data$mom_numcig)

#changing character variables to factors/ numeric
data$income <- as.numeric(data$income)

data <- data %>%
  mutate_if(is.character, factor)

#summarize missingness by column
miss_var <- as.data.frame(miss_var_summary(data)) 
miss_var$pct_miss <- round(miss_var$pct_miss, 2)

miss_var[miss_var$pct_miss > 25,] %>%  
  knitr::kable(caption = "Summary of Missing Data by Variable",
                                                    col.names = c("Variable",
                                                                  "Count",
                                                                  "Percentage")) %>%  
  kableExtra::kable_styling( latex_options = c("striped"),
                             stripe_color = "gray!15")

```



```{r}
#summarize missingness by row
pct_na_r <- round(rowSums(is.na(data)) / ncol(data) * 100,2)
row_na <- data.frame(parent_id = data$parent_id, pct_na = pct_na_r)
row_na <- row_na[row_na$pct_na > 50,] %>% arrange(desc(pct_na)) # participants w/ greater than 50% missing

row_na %>%  
  knitr::kable(caption = "Summary of Missing Data by Parent",
                                                    col.names = c("Parent ID",
                                                                  "Percentage missing")) %>%  
  kableExtra::kable_styling(latex_options = c("striped"),
                             stripe_color = "gray!15")

```


```{r}
#manually imputing values for SDP and PP
data$mom_smoke_32wk  <- ifelse(data$mom_smoke_16wk == "1=Yes" & 
                                 data$mom_smoke_22wk == "1=Yes",
                  "1=Yes", as.character(data$mom_smoke_32wk) )

data$mom_smoke_pp1 <- ifelse(data$mom_smoke_22wk == "1=Yes" & 
                               data$mom_smoke_32wk == "1=Yes" &
                               data$mom_smoke_pp2 == "1=Yes",
                  "1=Yes", as.character(data$mom_smoke_pp1) )
 
```


### Baseline Characteristics 

For this sample, 53.06% of parents were white and 26.53% of parents were Hispanic/Latino.
This sample did not include asian nor black parents. One parent reported being assigned
Male at birth. 30.61% of children were black, 30.61% were Hispanic/Latino, 
38.78% were white, and none identified as asian. Of the children, 26.53% were 
female, and 46.94% were male. The mean income was 44,424 with 20,000 and 66,250 being the 25th and 75th quartile. The max and min, respectively, for income are 265,000 
and 760. 54% of parents are employed full-time and 29% are unemployed. Also, 
37% of parents had some college  education and 24% have a 4-year degree. Note, these baseline characteristics included all parents (N = 49).



```{r}
#socioeconomic characteristics for sample
base_socioec <- data %>% select(income, employ, pedu) %>%
  tbl_summary(missing = "no") %>% 
   add_n(statistic = "{n_miss} ({p_miss}%)") %>%
  modify_header(n = "N Missing") %>%
  knitr::kable(caption = "Summary of Socioeconomic Status",
           col.names = c("Characteristic", "N Missing", "N = 49") ) %>%  
  kableExtra::kable_styling(latex_options = c("striped"),
                             stripe_color = "gray!15")

#max(data$income, na.rm = T)
#min(data$income, na.rm = T)

base_socioec
```



```{r}

#baseline characteristics for sample
baseline_cat <- data %>% select(psex,plang,pethnic, pasian,
pblack, pwhite, tsex, tethnic, tasian, tblack, twhite) %>%
  gather(variable, value) %>%  # Reshape to long format
  group_by(variable, value) %>%  
  summarise(count = n(),  # Count "Yes" and "No"
            percentage = round(n() / nrow(data) * 100,2)) %>% # get percentage
mutate(count_percentage = paste(count, " (", percentage, "%)", sep = ""))  # Combine count and percentage

baseline_cat$value <- as.factor(baseline_cat$value)

# Pivot the data to have "Yes"/ "No" columns
summary_df <- baseline_cat %>% select(variable, value, count_percentage) %>%
  pivot_wider(names_from = value, values_from = count_percentage)

summary_df<- summary_df %>% select("variable", "1", "0", "NA", "2")

summary_df[,] <- lapply(summary_df[,], function(x)ifelse(is.na(x), "0 (0%)", x))

base_char <- summary_df  %>% 
knitr::kable(caption = "Summary of Parent and Child Demographics",
                                                    col.names = c("Characteristic",
                           "Yes","No",  "NA", "Prefer Not To Answer")) %>%  
  kableExtra::kable_styling(latex_options = c("striped"),
                             stripe_color = "gray!15")

base_char
```


### Reported Smoking Status 

In the exploratory analysis for smoking behavior during pregnancy (See Fig 1 below),
we observe that the number of parents who reported smoking throughout different 
time points in their pregnancy remained
for the most part constant (near 22-27%). The percentage of parents who reported not
smoking during the first trimester decreased substantially from 75% to 62.5% 
during the third trimester. This observation is most likely caused by missing data 
which may have had an impact as we also observe that the first trimester had the least amount of missing data 0% (2.04% before removing parent id 53902), and the third 
trimester had the most missing data with 14.58%. In observing ETS exposure, overall
18% to 23% of parents reported their child being exposed to ETS at any of the time points 
surveyed. For this analysis, it was decided to focus on 12wk and 6 months postpartum
reported smoking due to the percentage of missing data 70.83% and 39.58% in reported smoking 
at postpartum visit 1 and 2 (after making assumptions on NA values), respectively. Using these variables may lead to biased results because of the underestimated
proportions of smoker and non-smoker parents at those visits as a result of missing data.


```{r, include = FALSE}
#exclusion criteria - missing all SDP/postpartum/ETS AND cotinine measurements
for (i in 1:nrow(data))
{
    if (sum(is.na(data[i,22:24])) == 3 &  sum(is.na(data[i,25:28])) == 4 & 
        sum(is.na(data[i,37:42])) == 6)
        
    {
      if(
        sum(is.na(data[i,29:31])) == 3)
        print("") #print(i) ### --- parent 53902
    }
}
#remove parent
data <- data[-c(38),] 

```

```{r}
##### This code section focuses on SDP variables #####
#summary of yes, no, na
t1 <- table(data$mom_smoke_16wk , useNA = "always")
t2 <- table(data$mom_smoke_22wk, useNA = "always")
t3 <- table(data$mom_smoke_32wk, useNA = "always")
#get percentages
df1 <- t(as.data.frame(round(prop.table(t1) * 100,2)))
df2 <- t(as.data.frame(round(prop.table(t2) * 100,2)))
df3 <-  t(as.data.frame(round(prop.table(t3) * 100,2)))
#create df containing all data
final <- rbind((df1)[2,],df2[2,],df3[2,])
colnames(final) <- c("Yes", "No", "NA")
time <- data.frame(Time = c("16wk", "22wk", "32wk"))
final <- cbind(time, final)
final$Yes <- as.numeric(final$Yes)
final$No <- as.numeric(final$No)
final$"NA" <- as.numeric(final$"NA")
#change to long format for plotting
final2 <- final %>% pivot_longer(cols = 'Yes':'NA',
                               names_to = "Category",
                               values_to = "Per")
#plot bar chart
SDP_plot <- ggplot(final2, aes(x = Time, y = Per, fill = Category)) +
  geom_col() +
  labs(y = "Percentage", title = "Parent Smoking Behavior During Pregnancy") +
   geom_text(aes(label = Per), position = position_stack(reverse = F,vjust = 0.25), size = 3)

##### This code section focuses on smoking PP variables #####
#summary of yes, no, na
t1 <- table(data$mom_smoke_pp1 , useNA = "always")
t2 <- table(data$mom_smoke_pp2, useNA = "always")
t3 <- table(data$mom_smoke_pp12wk, useNA = "always")
t4 <- table(data$mom_smoke_pp6mo, useNA = "always")
#get percentages
df1 <- t(as.data.frame(round(prop.table(t1) * 100,2)))
df2 <- t(as.data.frame(round(prop.table(t2) * 100,2)))
df3 <-  t(as.data.frame(round(prop.table(t3) * 100,2)))
df4 <-  t(as.data.frame(round(prop.table(t4) * 100,2)))
#create df containing all data
final <- rbind((df1)[2,],df2[2,],df3[2,],df4[2,])
colnames(final) <- c("Yes", "No", "NA")
time <- data.frame(Time = c("Postpartum V1", "Postpartum V2", "12wk postpartum",
                            "6mo postpartum"))
final <- cbind(time, final)
final$Yes <- as.numeric(final$Yes)
final$No <- as.numeric(final$No)
final$"NA" <- as.numeric(final$"NA")
#change to long format for plotting
final2 <- final %>% pivot_longer(cols = 'Yes':'NA',
                               names_to = "Category",
                               values_to = "Per")
#plot bar chart
pp_plot <- ggplot(final2, aes(x = Time, y = Per, fill = Category)) +
  geom_col() +
  labs(y = "Percentage", title = "Parent Smoking Behavior Postpartum") +
   geom_text(aes(label = Per), position = position_stack(reverse = F,vjust = 0.25), size = 3)

##### This code section focuses on ETS variables #####
#summary of yes, no, na
t1 <- table(data$smoke_exposure_6mo , useNA = "always")
t2 <- table(data$smoke_exposure_12mo, useNA = "always")
t3 <- table(data$smoke_exposure_2yr, useNA = "always")
t4 <- table(data$smoke_exposure_3yr, useNA = "always")
t5 <- table(data$smoke_exposure_4yr, useNA = "always")
t6 <- table(data$smoke_exposure_5yr, useNA = "always")

#get percentages
df1 <- t(as.data.frame(round(prop.table(t1) * 100,2)))
df2 <- t(as.data.frame(round(prop.table(t2) * 100,2)))
df3 <-  t(as.data.frame(round(prop.table(t3) * 100,2)))
df4 <-  t(as.data.frame(round(prop.table(t4) * 100,2)))
df5 <-  t(as.data.frame(round(prop.table(t5) * 100,2)))
df6 <-  t(as.data.frame(round(prop.table(t6) * 100,2)))

#create df containing all data
final <- rbind((df1)[2,],df2[2,],df3[2,],df4[2,], df5[2,], df6[2,])
colnames(final) <- c("No", "Yes", "NA")
time <- data.frame(Time = c("ETS 6mo", "ETS 12mo", 
                            "ETS 2yr","ETS 3yr","ETS 4yr","ETS 5yr"))
final <- cbind(time, final)
final$Yes <- as.numeric(final$Yes)
final$No <- as.numeric(final$No)
final$"NA" <- as.numeric(final$"NA")
#change to long format for plotting
final2 <- final %>% pivot_longer(cols = 'No':'NA',
                               names_to = "Category",
                               values_to = "Per")

final2$Time <- factor(final2$Time, levels=c( "ETS 6mo",
  "ETS 12mo" ,"ETS 2yr",  "ETS 3yr",  "ETS 4yr",  "ETS 5yr"))

#plot line chart
ETS_plot <- ggplot(final2, aes(x = Time, y = Per,fill = Category)) +
  geom_col() +
  labs(y = "Percentage", title = "Child Environmental Tobacco Exposure") +
  geom_text(aes(label = Per), position = position_stack(reverse = F,vjust = 0.25), size =3)
  

```


```{r fig.width=8,fig.height=8}
#View Plots of yes/no/na by smoking at different times
grid.arrange(SDP_plot, pp_plot, ETS_plot, heights = c(4,4,4))
```




### Composite Scores, Correlations and Possible Associations

From the SDP data, a cumulative SDP (Smoking During Pregnancy) variable was created.
If a parent reported smoking at 16,22, and 32 weeks, then the cumulative sum and
grouping was 3. A overall average was also created for smoking postpartum at 12 weeks and 
up to 6months after. A percentage score for ETS exposure from 6months to 5 years was also created, and in the tables is reported as the total years out of 5 for which a parent responded 'yes' to their child being exposed to ETS. 

The data also includes various variables that measure similar characteristics which 
can be classified relating to self-regulation or externalizing. In considering 
fitting a linear model, it may be of interest to create composite scores through 
normalizing variables if they scores measuring  similar characteristics but reported on different scales. The normalized scores can then be averaged. These composite scores will still be representative of the degree or severity of the responses. 

In observing the correlation values between SR/ EXT variables (see correlation table below),
there are a few variable pairs that have strong positive or negative correlations near 0.7 or -0.7. For strong correlations between hyperactive and inattentive pairs or inattentive/ hyperactive and parent's response on child externalizing, this is good to see as we should
expect these variables to correlate or not contradict.

Another observation being child hyperactivity and parent's self response on expressive suppression pairs have a 0.7888 correlation. Moreover, the parent's response on their child's attention problem also has a 0.7898 correlation with the parent's self response on expressive suppression. The negative correlation observed -0.7325 corresponds to a child's self internalizing score and the parent's internalizing 
score on self. It is possible that a parent's externalizing or internalizing behaviors may
affect the child's self-regulation or externalizing scores. 


```{r}
#creating composite SDP/PP/ETS variable
data$cumSDP <- rowSums(data[,c("mom_smoke_16wk",
                            "mom_smoke_22wk", 
                            "mom_smoke_32wk")] == "1=Yes",
                    na.rm = T)

data$cumPP <- (rowSums(data[,c("mom_smoke_pp12wk",
                             "mom_smoke_pp6mo")] == "1=Yes",
                    na.rm = T))


data$avgETS <- (((rowSums(data[,c("smoke_exposure_6mo",
                            "smoke_exposure_12mo")]  == 1, na.rm = T))/2)
   + (rowSums(data[,c( "smoke_exposure_2yr", "smoke_exposure_3yr", 
                       "smoke_exposure_4yr", "smoke_exposure_5yr")]  == 1,
              na.rm = T)))/5


```



#### Examining cotinine measurement data

The data also included cotinine measurements at 34 weeks gestation and 6 months
postpartum for the parent and at 6 months postpartum for child. We can compare
the mean values for these results across SDP.

In the line plot below, we observe that for parents who reported not smoking during any
trimester in their pregnancy, their cotinine levels at 34 weeks gestation were near 0.
For a few parents, their cotinine levels increased at 6 months postpartum. This could be
due to parents resuming smoking after the giving birth. Parents who reported smoking
their entire pregnancy have significantly higher cotinine measures compared to parents who 
reported not smoking their entire pregnancy or who only reported smoking one trimester.
Some of these values are drastically high, reaching up to a maximum of 878.96. It would be
good to consider parent smoking behavior before pregnancy since if they continued smoking
during pregnancy it may have an aggregated affect.



```{r fig.width=5,fig.height=3}
#pivot data for plotting purposes
data1 <- data %>% select(parent_id, avgETS,cumPP, cumSDP, 
                         cotimean_34wk, cotimean_pp6mo) %>%
  pivot_longer(cols = starts_with("cotimean"), names_to = "time", values_to = "measurement")

#re leveling data
data1$time <- as.factor(data1$time)
data1$cumSDP <- as.factor(data1$cumSDP)
data1$cumPP <- as.factor(data1$cumPP)

levels(data1$cumSDP) <- c("No trimesters","One trimester","Entire pregnancy")
levels(data1$cumPP) <- c("No","3 mo. postpartum","6 mo. postpart.")

#spaghetti plot of cotinine measurements stratified by cumSDP
ggplot(data1, aes(x = time, y = measurement, group = parent_id)) +
  geom_line() +
  facet_grid(cols = vars(cumSDP)) +
  theme_bw() +
  labs(x = "Time", y = "Parent Cotinine level", main = "Change in parent cotinine levels
       at different time points") +
  scale_x_discrete(labels = c("3rd trim.", "6mo. postpar."))

#max(data$cotimean_pp6mo , na.rm = T) 

#spaghetti plot of cotinine measurements stratified by postpartum smoking
#ggplot(data1, aes(x = time, y = measurement, group = parent_id)) +
#  geom_line() +
#  facet_grid(cols = vars(cumPP)) +
#  theme_bw() +
#  labs(x = "Time", y = "Cotinine level", main = "Change in parent cotinine levels
#       at different time points") +
#  scale_x_discrete(labels = c("Third trim.", "6 months postpart."))

```


Additionally, from the dot plot below we observe the reported number of days a parent smoked in the past 30 days and the number of cigarettes usually smoked per day. A majority of parents reported not smoking. One parent reported smoking less than 10 days in the past 30 days, but smoked over 20 cigarettes per day. Parents who reported smoking 30 days, reported usually smoking near 5-10 cigarettes a day. Here we also see a a value outside the possible 0-30 day range.


```{r fig.width=3,fig.height=3}
#plot of parent smoking behavior
ggplot(data, aes(x = momcig, y = mom_numcig)) +
geom_point(alpha = 0.25, size = 2.8) + theme_light() +
  labs(x = "Number of days", y = "Number of Cigarettes Smoked",
       title = "Parent smoking behavior")   #past 30 days
```





In the plot of parent and child cotinine levels, we see some correlation between
parent cotinine level at 34 weeks and 6 months postpartum with child cotinine level at 6 months postpartum. In further exploring this, using Pearson's correlation, parent and
child cotinine levels at 6 months have a 0.627 correlation with p-value = 2.514e-05 and 36 degrees of freedom. The plot also shows a potential outlier. 


```{r}

#plot of cotinine measures
par(mfrow=c(1,2))
plot(data$cotimean_34wk, data$cotimean_pp6mo_baby, 
     xlab = "Parent cotinine level in third trimester",
     ylab = "Child cotinine level (6 mo. postpart.)")
plot(data$cotimean_pp6mo,data$cotimean_pp6mo_baby,
      xlab = "Parent cotinine level 6 months postpartum", ylab = "Child cotinine level (6 mo. postpart.)")
mtext("Parent and Child Cotinine Levels", side = 3, line = -3, outer = T)

#perform correlation test
cor.test(data$cotimean_34wk, data$cotimean_pp6mo_baby)
cor.test(data$cotimean_pp6mo,data$cotimean_pp6mo_baby)

```



```{r}
### This section focuses on correlation ###

#select columns of interest relating to SR/EXT
columns_select <- c("childasd", "swan_inattentive","swan_hyperactive" ,          "bpm_att_p","bpm_ext_p","bpm_int_p", "bpm_att_a", "bpm_ext_a",                 
 "bpm_int_a", "erq_cog_a" , "erq_exp_a" , "bpm_att" , "bpm_ext", "bpm_int",                 
 "erq_cog"   , "erq_exp" )

#get correlation matrix          
subset <- data[,columns_select] %>% select_if(is.numeric) 
correlation_matrix<- as.data.frame(cor(subset,use = "complete.obs"))
correlation_matrix<- round(cor(subset,use = "complete.obs"),4)

#select correlation values > 0.7
new_matrix<- as.data.frame(as.table(correlation_matrix))
new_matrix<- new_matrix %>% filter(Freq != 1 & abs(Freq) > 0.7)
new_matrix <- new_matrix[-c(2,3,4,9,13,14),] 
colnames(new_matrix) <- c("Variable 1", "Variable 2", "Correlation")
new_matrix %>% knitr::kable()


### This code section focuses on normalizing and creating composite scores ###
#standardized_A <- (data$erq_exp - mean(data$erq_exp, na.rm = T)) / sd(data$erq_exp,           na.rm = T)
#normalized_B <- (data$erq_exp - min(data$erq_exp, na.rm = T)) / (max(data$erq_exp, na.rm = T) - min(data$erq_exp, na.rm = T))
#quantile(normalized_B, probs = c(0.25, 0.75), na.rm = T)
```

### Aim 1: Examine effects of SDP/ETS on adolescent self-regulation, substance use, and externalizing. 



### Self-Regulation and Externalizing

The summary tables below include both score responses for child on self and parent on child and swan scores. It should be expected that their responses match closely, however, the resulting tables show that children tended to report higher scores and parents reported lower scoring.
This shows a potential disconnect between parent and child. A possible next step would be to consider whether parental knowledge has any impact on these differences.

In table 5, parents who reported smoking during all trimesters (N = 11) had higher mean child reported scores for attention and externalizing (relating to externalizing). Parents who reported smoking during any one trimesters (N = 3) had higher mean values for their child internalizing, cognitive reappraisal,
and expressive suppression (relating to self-regulation). It is also important to note that for swan hyperactive ad inattentive scores, there is a large variability in scores (as seen in the standard deviation, 8.45 and 7.93 respectively). Overall, there are a few low but not significant p-values (for child/ parent reported attention problems and child reported expressive suppression). 



```{r}
### summarizing SR and EXT variables ###
data1 <- data
#re leveling data
data1$cumSDP <- as.factor(data1$cumSDP)
data1$cumPP <- as.factor(data1$cumPP)
levels(data1$cumSDP) <- c("0/3 tri.","1/3 tri.", "3/3 tri.")
levels(data1$cumPP) <- c("No","For Three months postpartum","All 6 months postpartum")


### stratified by cum SDP
behav_SDP <- data1  %>% select(cumSDP, bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) %>%
  tbl_summary(by = cumSDP, missing = "no",
                     type = list (c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ "continuous2",
                 c(cumSDP) ~ "categorical"),
                     statistic = all_continuous() ~ c("{mean} ({sd})",
                           "{N_miss} ({p_miss}%)"),
              digits = list(c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ c(2, 2))) %>%
        add_p() %>% add_ci() %>% knitr::kable(caption = "Summary of Self-Regulation and Externalizing Behavior by Cumulative SDP
            Exposure") %>%  kableExtra::kable_styling(latex_options = c("striped"),
                             stripe_color = "gray!15")


### stratified by smoking postpartum indicator cum score ###
behav_cumPP <- data1  %>% select(cumPP, bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) %>%
  tbl_summary(by = cumPP, missing = "no",
                type = list(c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ "continuous2",
                 c(cumPP) ~ "categorical"),
                     statistic = all_continuous() ~ c("{mean} ({sd})",
                           "{N_miss} ({p_miss}%)"),
            digits = list(c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ c(2, 2))) %>%
  add_p() %>%
knitr::kable(caption = "Summary of Self-Regulation and Externalizing Behavior") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")



behav_pp12wk <- data1  %>% select(mom_smoke_pp12wk, bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) %>%
  tbl_summary( by = mom_smoke_pp12wk, missing = "no",
                     type = c( bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ "continuous2",
                     statistic = all_continuous() ~ c("{mean} ({sd})","{N_miss} ({p_miss}%)"),
            digits = list(c( bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p,erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ c(2, 2)))%>% add_p() %>% add_ci() %>% knitr::kable(caption = "Summary of Self-Regulation and Externalizing Behavior by 12 weeks Postpartum Smoke Exposure") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")


behav_pp6mo <- data1  %>% select(mom_smoke_pp6mo, bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p) %>%
  tbl_summary( by = mom_smoke_pp6mo,
                     type = c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p) ~ "continuous",
                     statistic = all_continuous() ~ "{mean} ({sd})",
            digits = list(c(bpm_att, bpm_ext, bpm_int, 
                 bpm_att_p, bpm_ext_p, bpm_int_p) ~ c(2, 2))) %>% add_p()

### stratified by ETS avg score ###

#releveling data
data1$avgETS <- as.factor(data1$avgETS)
levels(data1$avgETS) <- c("0","0.5", "2.5", "3", "3.5", "4", "4.5", "5")
 
behav_ETS <- data1  %>% select(avgETS, bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) %>%
  tbl_summary(by = avgETS, missing = "no",
                     type = c( bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ "continuous2",
                     statistic = all_continuous() ~ c("{mean} ({sd})",
                           "{N_miss} ({p_miss}%)"),
              digits = list(c( bpm_att, bpm_att_p, bpm_ext,
                 bpm_ext_p,  bpm_int, bpm_int_p, erq_cog, erq_exp,
                 swan_hyperactive, swan_inattentive) ~ c(2, 2))) %>% add_p() %>%  knitr::kable(caption = "Summary of Self-Regulation and Externalizing Behavior by ETS exposure") %>% kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15", full_width = F)


```



```{r fig.width=12,fig.height=10}
behav_SDP

```




In table 6, stratifying by child postpartum smoke exposure at 12 weeks, parents who responded
yes showed higher mean values for all child self-regulation and externalizing scores compared to parents who responded no. There was a statistically significant difference in scores for attention problems reported by both the child and parent and child internalizing problems reported by the parent.  


```{r}
behav_pp12wk

```





```{r}

behav_ETS

```






In table 7, it is noted that due to the small sample size, there was one parent who 
reported child's ETS for 2.5, 3, 3.5 and 4.5 years. In general, parents who responded
no for ETS exposure for all 5 years had the lowest mean scores for all categories with a few exceptions. Also, the higher mean values did not occur for parents who responded yes for 
ETS for all 5 years, but for children exposed for 0.5 to 3.5 range. Any added 0.5 corresponds to 6 months or 12 months reported postpartum smoking. It would be of interest to explore this further. In the tables not included in this report, we see a few significant values when comparing outcomes stratified by postpartum smoking (see code behav_pp12wk and behav_pp6mo).



#### Substance Use

For this aim, we also summarize reported substance use stratified by cumulative SDP and average postpartum smoking.

Table 8 and 9 shows substance use stratified by SDP exposure. The only statistically significant value observed is for number of alcohol, however for these results, it is difficult
to tell due to the small sample sizes, hence the large confidence intervals shows for the
true proportions that fall within that category. It would be of interest to further explore how
pospartum exposure affects substance use based on table 10.

```{r}

#re leveling data
#data1$avgETS <- as.factor(data1$avgETS)
#levels(data1$avgETS) <- c("0","0.5", "2.5", "3", "3.5", "4", "4.5", "5")
 
#stratified by SDP
SU_sdp <- data1  %>% select(cumSDP, cig_ever, e_cig_ever, alc_ever, mj_ever) %>%
  tbl_summary(by = cumSDP, missing_text = "Missing",
                     statistic = all_continuous() ~ "{n} ({p}%)",
                         #all_continuous() ~ "{mean} ({sd})",
              digits = list(c(cig_ever, e_cig_ever, alc_ever,
                              mj_ever) ~ c(2, 2))) %>% add_p()%>% add_ci() %>% knitr::kable(caption = "Summary of Substance Use by cumulative SDP Exposure") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")

SU_sdp2 <- data1  %>%  select(cumSDP, num_cigs_30,
                              num_e_cigs_30, num_alc_30, num_mj_30) %>%
                  tbl_summary(by = cumSDP, missing_text = "Missing",
                     type = c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ "continuous",
                     statistic = all_continuous() ~ "{mean} ({sd})",
                         #all_continuous() ~ "{mean} ({sd})",
              digits = list(c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ c(2, 2))) %>% add_p()%>% 
  knitr::kable(caption = "Summary of Substance Use by cumulative SDP Exposure") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")

#stratified by cum PP
SU_pp <- data  %>% select(cumPP, cig_ever, num_cigs_30, e_cig_ever, num_e_cigs_30, alc_ever,
                 num_alc_30, mj_ever, num_mj_30) %>%
  tbl_summary(by = cumPP, missing_text = "Missing",
                     type = c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ "continuous",
                     statistic = all_continuous() ~ "{mean} ({sd})",
                         #all_continuous() ~ "{mean} ({sd})",
              digits = list(c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ c(2, 2))) %>%  add_p() %>%  knitr::kable(caption = "Summary of Substance Use by Postpartum Exposure") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")

#stratified by avg ETS
SU_ets <- data1  %>% select(avgETS, cig_ever, num_cigs_30, e_cig_ever, num_e_cigs_30, alc_ever, num_alc_30, mj_ever, num_mj_30) %>%
  tbl_summary(by = avgETS, 
                     type = c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ "continuous",
                     statistic = all_continuous() ~ "{mean} ({sd})",
                         #all_continuous() ~ "{mean} ({sd})",
              digits = list(c(num_cigs_30, num_e_cigs_30, num_alc_30,
                              num_mj_30) ~ c(2, 2))) %>% add_p() %>% 
  knitr::kable(caption = "Summary of Substance Use by  Exposure") %>%  
  kableExtra::kable_styling(latex_options = c("striped"), stripe_color = "gray!15")


```



```{r}
SU_sdp
SU_sdp2
SU_pp
#SU_ets - not shown in report due to small sample sizes across categories

```



\newpage

## Code Appendix:

```{r get-labels, echo=FALSE}
labs = knitr::all_labels()
labs = setdiff(labs, c("get-labels"))
```

```{r all-code, ref.label=labs, echo=TRUE, eval=FALSE}
```


